---
title: "Gender Clumping in Gov 1006"
author: "David Kane"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

Do students clump by gender in Gov 1006? 

Let's ignore the table with all graduate students, since this will never clump. Let's ignore the table with two graduate students. (It is always a good idea to make the initial version of a simulation as simple as possible.) There are 30 remaining students: 16 men and 14 women. Our null model is that they sit down randomly. We can ignore the last two students since they are, by assumption, sitting at the table with 2 graduate students. This leaves us with 28 students (chosen randomly from the 30 undergraduates) sitting at 7 tables of 4 students each.

Define a "clump" as a table of four, all of the same gender. (Other definitions of clumping are possible. One aspect of data science is going from a verbal description to a mathematical formulation.) In class on February 12th, there were three such tables. Is that surprising? Is it evidence of clumping? If students sat randomly, how unusual would it be to see three single-gender tables?


New question (2/19) - how likely to see female only clumps?

```{r tools}
# Before we run the simulation, we need the basic data and some simple
# functions. Removing the graduate students, there are 16 men and 14 women in
# the class. Let's use a vector to represent that information. Implicitly, the
# first four values in the vector sit at table 1, the second four sit at table 2
# and so on.

class30 <- c(rep("M", 16), rep("F", 14))

seat_a_class <- function(){
  
  # A class seating is defined as 28 of the 30 students in a vector. We only
  # need 28 because the last two, by assumption, sit at the table with Cian/Alex
  # and, therefore, don't matter to our calculations. The function could be a
  # one-line --- list(sample(class30)[1:28]) --- but I like to show the individual
  # steps.
  
  # I think that returning a list is necessary because I am going to assign the
  # result as a list-column!!
  
  x <- sample(class30)
  x <- x[1:28]
  list(x)
}

calc_clumped_tables <- function(seating){
  
  # Given a seating --- which, again, is just a vector of length 28 with Ms and
  # Fs in random order --- how many clumped tables are there? We will define a
  # clumped table as one that is all M or all F. As discussed in class, there
  # are other definitions of "clumping" that we could use.
  
  # Good functions test to see if their assumptions are met. See the chapter in
  # the new 1005 textbook for discussion:
  # https://davidkane9.github.io/PPBDS/6-functions.html
  
  stopifnot(is.list(seating))
  plan <- unlist(seating)
  
  # MOLLY: use stopifnot() to assert things and tell the function to stop if any of these are not true
  # MOLLY: in this case we want the insert to be a list, with 28 ppl, that are M or F only
  # MOLLY: this will stop nonsense results from being created and instead throw an error
  
  stopifnot(length(plan) == 28)
  stopifnot(all(plan %in% c("M", "F")))

  # It is not obvious how to do this elegantly. So, I just hack it! (Stole this
  # approach from Beau, I think.) I grab out each table, one at a time, knowing
  # that I have arbitrarily decided that seats 1 through 4 are Table 1, seats 5
  # through 8 are Table 2 and so on. Then, I create a logical vector reporting
  # whether or the person at each seat is Female. Then, I sum up this vector. If
  # it is all Male, the sum will be zero. It it is all Female, the sum will be
  # four. Check that, and sum of the results for each table. Obviously, you need
  # to do a fair bit of fussing to make sure this is working. Test cases would
  # be good!
  
  clumps <- 
    (sum(plan[1:4] == "F") %in% c(0, 4)) +
    (sum(plan[5:8] == "F") %in% c(0, 4)) +
    (sum(plan[9:12] == "F") %in% c(0, 4)) +
    (sum(plan[13:16] == "F") %in% c(0, 4)) +
    (sum(plan[17:20] == "F") %in% c(0, 4)) +
    (sum(plan[21:24] == "F") %in% c(0, 4)) +
    (sum(plan[25:28] == "F") %in% c(0, 4))
  
  return(clumps)
}



calc_female_clumped_tables <- function(seating){
  
  stopifnot(is.list(seating))
  plan <- unlist(seating)
  
  stopifnot(length(plan) == 28)
  stopifnot(all(plan %in% c("M", "F")))

  clumps <- 
    (sum(plan[1:4] == "F") == 4) +
    (sum(plan[5:8] == "F") == 4) +
    (sum(plan[9:12] == "F") == 4) +
    (sum(plan[13:16] == "F") == 4) +
    (sum(plan[17:20] == "F") == 4) +
    (sum(plan[21:24] == "F") == 4) +
    (sum(plan[25:28] == "F") == 4)
  
  return(clumps)
}


```

```{r sim, cache=TRUE}
# Those are all the tools we need. Many/most of my simulations use the tricks we
# learned in Gov 1005: list-colums and map_* functions are your friends!
# https://davidkane9.github.io/PPBDS/6-functions.html#list-columns-and-map_-functions

# I don't really need cache=TRUE since the code does not take *that* long to
# run. I just wanted to remind you of that trick.

sim <- 
  tibble(replication = 1:10000) %>% 
  mutate(seating_plan = map(replication, ~ seat_a_class())) %>% 
  mutate(clumps = map_int(seating_plan, ~ calc_clumped_tables(.)))

# map function says: take first thing (usually a column of a table) - and run this function on that - and store the result in each row of the tibble

# use str() function to look at the results of a column from sim 

female_sim <-
  tibble(replication = 1:10000) %>% 
  mutate(seating_plan = map(replication, ~ seat_a_class())) %>% 
  mutate(clumps = map_int(seating_plan, ~ calc_female_clumped_tables(.)))
  
 
```

```{r graphics}
# Once we have the data, plotting is straightforward. The ..prop.. trick
# required some Googling. Of course, I could just have calculated the
# percentages by hand.

sim %>% 
  ggplot() + 
    geom_bar(mapping = aes(x = clumps, y = ..prop..), 
             width = 0.1) +
    labs(title = "Gender Clumping in Gov 1006",
         subtitle = "Having three single-gender tables is surprising",
         caption = "Data Collected in Gov 1006 on 2020-02-12",
         x = "Number of Single-Gender Tables out of Seven",
         y = "Percentage of Simulations out of 10,000")

female_sim %>% 
  ggplot() + 
    geom_bar(mapping = aes(x = clumps, y = ..prop..), 
             width = 0.1) +
    labs(title = "Female Clumping in Gov 1006",
         subtitle = "Having two female-only tables is surprising",
         caption = "Data Collected in Gov 1006 on 2020-02-12",
         x = "Number of Female-Only Tables out of Seven",
         y = "Percentage of Simulations out of 10,000")
```

Given that we observed a value of 3 which occurs ~3% of the time--less than 5%--our observation fulfills the traditional definition for statistical significance. Thus, what we observed in the world is statsitically significant from the null hypothesis, so in that case where 3 tables were clumped, we can reject the null hypothesis and say in that case, the students DO clump by gender (not sitting randomly as the null hypothesis stated).

p-value - the probaility of observing something at least as extreme as the ovserved test statistic 
  - p-value is a summary of deviation of the null hypothesis
  - in this case, our observed test statistic would be our real world observation of 3 clumped tables
    - test statistic overall is number of tables w/ single gender
    - observed test statistic was 3 
  - in this case - our p-value would be count(clumps > 3) -- at least as extreme as the observed test statistic -- where we get 3.07% which is less than 5% and can reject the null
  - if you are greater than 5%, you would summarize what you get and interperet the value
  
```{r making fake grades}

ppl_in_class <- 80
grades <- rnorm(ppl_in_class, 75, 10)

noise1 <- rnorm(ppl_in_class, 0, 10)
noise2 <- rnorm(ppl_in_class, 0, 10)

midterm <- grades + noise1
final <- grades + noise2

exams <- data.frame(midterm,final) %>%
  mutate(complainer = NA) %>%
  mutate(treatment = NA)

for(i in 1:nrow(exams)) {
  
  ifelse(exams$midterm < 75, 
         exams$complainer[i] <- rbinom(1, 1, .67),
         exams$complainer[i] <- rbinom(1, 1, .33))
  
}

for(i in 1:nrow(exams)) {
  
  ifelse(exams$complainer == 1, 
         exams$treatment[i] <- rbinom(1, 1, .5),
         exams$treatment[i] <- 0)
  
}

```
  

